---
title: "Lab 4 Rowe"
author: "Joshua Rowe"
format:
  html:
    code-fold: true
editor: visual
execute: 
  warning: false
  message: false
---

## Setup

```{r}
#| warning: false
#| message: false

library(tidyverse)

number <- function(x) format(x, scientific = FALSE, big.mark = ",", digits = 3)

avocado <- read_csv(here::here("supporting_artifacts", "continued_learning",
                               "Data", "avocado.csv"))
```

## Data Cleaning

```{r Question 2}

## Question 2

regions <- avocado |> 
  distinct(region)

# Note: NewYork was assumed to be the city of new york because the magnitude of its "total volume" is vastly smaller than that of the state of California, a similarly sized state to New York, and is comparable in size to that of Chicago, a similarly sized city

states <- data.frame(state = c("California", "SouthCarolina"))

geo_regions <- data.frame(geo_region = c("GreatLakes", "Northeast",
                                         "NorthernNewEngland", "Plains",
                                         "SouthCentral", "Southeast", "West",
                                         "WestTexNewMexico", "Midsouth"))

total_US <- data.frame(total_us = "TotalUS")

# Old code:
# metro_areas <- avocado |> 
#   anti_join(states, by = c("region" = "state"))
# 
# metro_areas <- metro_areas |> 
#   anti_join(geo_regions, by = c("region" = "geo_region"))
# 
# metro_areas <- metro_areas |> 
#   anti_join(total_US, by = c("region" = "total_us")) |> 
#   rename(Small_Medium = `4046`, 
#          Large = `4225`, 
#          Extra_Large = `4770`,
#          Total_Volume = `Total Volume`)

# Updated code:
metro_areas <- avocado |> 
  anti_join(states, by = c("region" = "state")) |> 
  anti_join(geo_regions, by = c("region" = "geo_region")) |> 
  anti_join(total_US, by = c("region" = "total_us")) |> 
  rename(Small_Medium = `4046`, 
         Large = `4225`, 
         Extra_Large = `4770`,
         Total_Volume = `Total Volume`)


```

```{r Question 3}

## Question 3

# Old code:
# p3_filter <- data.frame(year = 2017, type = "organic")
# 
# p3 <- metro_areas |> 
#   select(Small_Medium, region, year, type) |> 
#   semi_join(p3_filter, by = "year") |> 
#   semi_join(p3_filter, by = "type") |> 
#   group_by(region) |> 
#   summarize(Small_Medium = max(Small_Medium))
# 
# p3_max <- data.frame(Small_Medium = max(p3$Small_Medium))
# 
# p3 <- p3 |> 
#   semi_join(p3_max, by = "Small_Medium")

# Updated code:
p3 <- metro_areas |> 
  select(Small_Medium, region, year, type) |> 
  filter(year == 2017, type == "organic") |> 
  slice_max(Small_Medium)
  

```

3.  The metro area that sold the most small/medium organic Haas Avocados is `r p3$region[1]`.

```{r Question 6}

CA_regions = data.frame(region = c("LosAngeles", "SanDiego", "Sacramento",
                                   "SanFrancisco"))

# Old code:
# p6 <- metro_areas |> 
#   semi_join(CA_regions, by = "region") |> 
#   select(region, type, AveragePrice) |> 
#   mutate(region = as.factor(region),
#          type = as.factor(type)) |> 
#   group_by(region, type) |> 
#   summarize(AveragePrice = mean(AveragePrice))

# Updated code:
p6 <- metro_areas |> 
  semi_join(CA_regions, by = "region") |> 
  select(region, type, AveragePrice) |> 
  mutate(region = as.factor(region),
         type = as.factor(type)) |> 
  group_by(region, type) |> 
  summarize(AveragePrice = mean(AveragePrice)) |> 
  pivot_wider(names_from = type, values_from = AveragePrice) |> 
  mutate(diff_price = organic - conventional)


ggplot(data = p6) + 
  geom_point(mapping = aes(x = conventional, 
                           y = diff_price,
                           color = region)) + 
  labs(x = "Price of Conventional Avocados [$]", 
       y = "Conventional/Organic Price Diff [$]", 
       color = "Region")


```

The largest difference in organic vs. non-organic avocado prices is seen in the San Francisco region, which has an average price of \$`r number(p6$organic[4])` for organic avocados and \$`r number(p6$conventional[4])` for non-organic avocados (difference of \$`r number(p6$diff_price[4])`). The second largest difference in prices is found in the Sacramento region with an average organic price of \$`r number(p6$organic[2])` and an average non-organic price of \$`r number(p6$conventional[2])` (difference of \$`r number(p6$diff_price[2])`).
