---
title: "Lab 4 Rowe"
author: "Joshua Rowe"
format:
  html:
    code-fold: true
editor: visual
execute: 
  warning: false
  message: false
---

## Setup

```{r}
#| warning: false
#| message: false

library(tidyverse)

number <- function(x) format(x, scientific = FALSE, big.mark = ",", digits = 3)

avocado <- read_csv(here::here("Lab Assignments", "Lab 4", "avocado.csv"))

```

1.  The dataset contains information from Hass brand avocados that were scanned through store registers from multiple stores from 2013 to 2018. The data contains information on the date, store type, and region of the avocado sales; date is listed as the "date" datatype, year is numerical, and region is a character datatype. There are also three numeric columns with the PLU code of the avocado sold; PLU 4046 is "small/medium Hass avocados," PLU 4225 is "large Hass avocados," and PLU 4770 is "extra large Hass Avocados."

## Data Cleaning

```{r Question 2}

## Question 2

regions <- avocado |> 
  distinct(region)

# Note: NewYork was assumed to be the city of new york because the magnitude of its "total volume" is vastly smaller than that of the state of California, a similarly sized state to New York, and is comparable in size to that of Chicago, a similarly sized city

states <- data.frame(state = c("California", "SouthCarolina"))

geo_regions <- data.frame(geo_region = c("GreatLakes", "Northeast", "NorthernNewEngland", "Plains", "SouthCentral", "Southeast", "West", "WestTexNewMexico", "Midsouth"))

total_US <- data.frame(total_us = "TotalUS")

metro_areas <- avocado |> 
  anti_join(states, by = c("region" = "state")) |> 
  anti_join(geo_regions, by = c("region" = "geo_region")) |> 
  anti_join(total_US, by = c("region" = "total_us")) |> 
  rename(Small_Medium = `4046`, 
         Large = `4225`, 
         Extra_Large = `4770`,
         Total_Volume = `Total Volume`)


```

```{r Question 3}

## Question 3

p3 <- metro_areas |> 
  select(Small_Medium, region, year, type) |> 
  filter(year == 2017, type == "organic") |> 
  slice_max(Small_Medium)
  

```

3.  The metro area that sold the most small/medium organic Haas Avocados is `r p3$region[1]`.

```{r Question 4}

metro_areas <- metro_areas |> 
  separate(Date, into = c("Year", "Month", "Day")) |> 
  mutate(Year = as.numeric(Year), 
         Month = as.numeric(Month), 
         Day = as.numeric(Day))

p4 <- metro_areas |> 
  select(Total_Volume, Month) |> 
  slice_max(Total_Volume)


```

4.  The highest mean volume of avocados, over the years 2015-2018, was sold in February (month 2).

```{r Question 5}

p5 <- metro_areas |> 
  select(region, Total_Volume) |> 
  group_by(region) |> 
  summarize(Total_Volume = sum(Total_Volume)) |> 
  slice_max(Total_Volume, n = 5)

p5_regions <- metro_areas |> 
  select(region, Total_Volume) |> 
  semi_join(p5, by = "region") |> 
  mutate(region = as.factor(region))

ggplot(data = p5_regions, mapping = aes(x = region, y = Total_Volume)) + 
  geom_boxplot() +
  labs(x = "Region", y = "Total Volume [# sold]")

```

```{r Question 6}

CA_regions = data.frame(region = c("LosAngeles", "SanDiego", "Sacramento", "SanFrancisco"))

p6 <- metro_areas |> 
  semi_join(CA_regions, by = "region") |> 
  select(region, type, AveragePrice) |> 
  mutate(region = as.factor(region),
         type = as.factor(type)) |> 
  group_by(region, type) |> 
  summarize(AveragePrice = mean(AveragePrice)) |> 
  pivot_wider(names_from = type, values_from = AveragePrice) |> 
  mutate(diff_price = organic - conventional)


ggplot(data = p6) + 
  geom_point(mapping = aes(x = conventional, 
                           y = diff_price,
                           color = region)) + 
  labs(x = "Price of Conventional Avocados [$]", 
       y = "Conventional/Organic Price Diff [$]", 
       color = "Region")


```

The largest difference in organic vs. non-organic avocado prices is seen in the San Francisco region, which has an average price of \$`r number(p6$organic[4])` for organic avocados and \$`r number(p6$conventional[4])` for non-organic avocados (difference of \$`r number(p6$diff_price[4])`). The second largest difference in prices is found in the Sacramento region with an average organic price of \$`r number(p6$organic[2])` and an average non-organic price of \$`r number(p6$conventional[2])` (difference of \$`r number(p6$diff_price[2])`).

```{r Question 7}

p7 <- metro_areas |> 
  select(region, type, Small_Medium, Large, Extra_Large) |> 
  semi_join(CA_regions, by = "region") |> 
  pivot_longer(cols = c(Small_Medium, Large, Extra_Large), 
               names_to = "size",
               values_to = "number") |> 
  mutate(across(.cols = region:size, ~as.factor(.x)),
         size = factor(size, levels = c("Small_Medium", "Large", "Extra_Large"))) |> 
  group_by(region, type, size) |> 
  summarize(number = sum(number))

# NOTE: In order to get the plot below to look like the desired plot from the lab description, the factors had to be re-ordered using the factor() function from dplyr which was found at the link https://r-graph-gallery.com/267-reorder-a-variable-in-ggplot2.html.

# In order to prevent the overlapping text on the x-axis in the same way that was done in the figure from the lab description, I found the guide_axis() argument at https://datavizpyr.com/how-to-dodge-overlapping-text-on-x-axis-labels-in-ggplot2/ (below).

ggplot(data = p7, mapping = aes(x = region, 
                                y = number, 
                                fill = size)) +
  geom_col(position = "fill") + 
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + 
  facet_wrap(~type) + 
  labs(x = "Region of CA", 
       y = "Proportion of Mean Avocados Sold",
       fill = "Avocado Size") + 
  scale_fill_discrete(labels = c("Small/Medium", "Large", "Extra Large"))

# Help on "scale_fill_discrete" was found at https://www.statology.org/ggplot-legend-labels/#:~:text=You%20can%20use%20the%20following,label3'%2C%20...))

```
