---
title: "Lab 1"
author: "Joshua Rowe"
format:
  html:
    code-fold: true
    toc: true
    toc-title: "This is an example of a Table of Contents"
    toc-location: "body"
    editor: visual
execute: 
  echo: true
  error: true
---

# Cars Dataset

```{r}

summary(cars)

```

This dataset contains information about speed and stopping distance of cars. The data was collected in the 1920's. It has 50 rows and two variables.

# Plot of Pressure Dataset

```{r}
#| fig-align: "center"
 
# Initial Code:
# plot(pressure, xlab = "Temperature [\u00B0C]", ylab = "Pressure [mmHg]")
# title(main = "Vapor Pressure of Mercury as a Function of Temperature")

# Revised Code:
plot(pressure, xlab = "Temperature [\u00B0C]", ylab = "Pressure [mmHg]")
title(main = "Vapor Pressure of Mercury as a Function of Temperature", sub = "Figure 1-1")

```

Note: The code-chunk option `#| fig-align: "center"` was used to specify that the above plot would be center-aligned.

# Plots of Cars Dataset

```{r}
#| layout-ncol: 2

dist <- cars$dist

hist(dist, main = "Histogram of Stopping Distance", 
     xlab = "Stopping Distance [ft]",
     sub = "Figure 2-1")
plot(dist, cars$speed, main = "XY Plot of Car Speed vs. Stopping Distance", 
     xlab = "Stopping Distance [ft]", 
     ylab = "Speed [mph]", sub = "Figure 2-2")


```

Things to improve about the plots above:

1.  I would like to play around with the bin size of the histogram to see if I could make it more informative, or if the automatically selected bin size seems to be the best representation of the data.
2.  It would be cool to change some aspects of the scatterplot, such as the color and type of the data points!
